### 2017-06-20 Notes/timing_and_convergence_examples.Rmd
updates to timing comparisons

---
title: "detrendr examples"
output: pdf_document
---

## Some examples of baseline fitting with detrendr

```{r load, include=FALSE, echo=FALSE, cache=FALSE}
library(Matrix)
library(devtools)
library(roxygen2)
library(Rcpp)
load_all("../Code/detrendr")

```

```{r simulate}

simulate <- function(n, k){
  x <- seq(1/n, 1, length.out=n)
  f <- 2*(x + 2)^2 + 3*cos(3*pi*x)
  tau <- 5e4
  g1 <- 40*exp(-tau*(x-0.3)^2)
  g2 <- 30*exp(-tau/2*(x-0.58)^2)
  g4 <- 45*exp(-tau/7*(x-0.6)^2)
  g3 <- 37*exp(-tau*(x-0.63)^2)
  y <- f + g1 + g2 + g3 + g4 + rnorm(n)
  D <- get_Dk(n, k)
  M <- diag(n) + crossprod(D)
  cholM <- Matrix::chol(M)
  return(list(y=y, f=f, x=x, D=D, M=M, cholM=cholM))
}

compareTimeR <- function(n){
  simData <- simulate(n, k)
  lambda <- lambda0*n^k/factorial(k)
  y <- simData$y
  theta <- simData$y
  D <- simData$D
  eta <- matrix(D%*%theta)
  M <- simData$M
  cholM <- simData$cholM
  spingarn_multi_step_R(theta, eta, y, D, M, lambda, tau, step, 500)
}

compareTimeC <- function(n){
  simData <- simulate(n, k)
  lambda <- lambda0*n^k/factorial(k)
  y <- simData$y
  theta <- simData$y
  D <- simData$D
  eta <- matrix(D%*%theta)
  M <- simData$M
  cholM <- simData$cholM
  spingarn_multi_step(theta, eta, y, D, M, lambda, tau, step, 500)
}
```

## Current timing experiments
```{r timing, include=TRUE, echo=TRUE, cache=TRUE}
nSeq <- seq(200, 3000, 200)
timing <- data.frame(n=nSeq, R_mean = NA, R_lq = NA, R_uq = NA,
                             C_mean=NA, C_lq=NA, C_uq = NA)
tau <- 0.01
lambda0 <- 1e-8
step <- 1
k <- 3

for (i in 1:length(nSeq)){

  set.seed(38912498)
  Rbench <- microbenchmark(compareTimeR(nSeq[i]), times=100)
  set.seed(38912498)
  Cbench <- microbenchmark(compareTimeC(nSeq[i]), times=100)

  timing$R_mean[i] <- median(Rbench$time*1e-9)
  timing$R_lq[i] <- quantile(Rbench$time*1e-9, .25)
  timing$R_uq[i] <- quantile(Rbench$time*1e-9, .75)
  timing$C_mean[i] <- median(Cbench$time*1e-9)
  timing$C_lq[i] <- quantile(Cbench$time*1e-9, .25)
  timing$C_uq[i] <- quantile(Cbench$time*1e-9, .75)
  print(i)
}

library(microbenchmark)

plot(C_mean~n, timing, type="l", col="blue", ylab="Elapsed Time (s)' per 500 iterations", main = "R code is in black, C in blue")
lines(C_lq~n, timing, col="blue")
lines(C_uq~n, timing, col="blue")
lines(R_mean~n, timing, type="l")
```

## Convergence when n=100, lambda = 10
```{r convergence, include=TRUE, echo=TRUE, cache=TRUE}

n <- 500
simData <- simulate(n, 3)

y <- simData$y
theta <- simData$y
D <- simData$D
eta <- matrix(D%*%theta)
M <- simData$M
cholM <- simData$cholM
tau <- 0.05
lambda <- 10
step <- 1

max_iter <- 70000
rerr <- double(max_iter-1)
THx <- matrix(NA, n, max_iter)
for (iter in 1:max_iter) {
  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
  theta <- one_step[[1]]
  eta <- one_step[[2]]
  THx[,iter] <- prox_f1(theta, y, tau)
  if (iter > 1){
    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
  }
}
theta_last <- prox_f1(theta, y, tau)

par(mfrow=c(2,1), mar=c(3,4,1,1))
plot(log(rerr), type="l")

plot(simData$x,simData$f,type='l',col='blue',
ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
points(simData$x,y, cex=.6)
lines(simData$x,theta_last,col='red', lwd=3)

```

## Convergence when n=1000, lambda = 100
```{r convergence2, include=TRUE, echo=FALSE, cache=TRUE}

n <- 1000
simData <- simulate(n, 3)

y <- simData$y
theta <- simData$y
D <- simData$D
eta <- matrix(D%*%theta)
M <- simData$M
cholM <- simData$cholM
tau <- 0.05
lambda <- 100
step <- 1

max_iter <- 70000
rerr <- double(max_iter-1)
THx <- matrix(NA, n, max_iter)
for (iter in 1:max_iter) {
  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
  theta <- one_step[[1]]
  eta <- one_step[[2]]
  THx[,iter] <- prox_f1(theta, y, tau)
  if (iter > 1){
    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
  }
}
theta_last <- prox_f1(theta, y, tau)

par(mfrow=c(2,1), mar=c(3,4,1,1))
plot(log(rerr), type="l")

plot(simData$x,simData$f,type='l',col='blue',
ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
points(simData$x,y, cex=.6)
lines(simData$x,theta_last,col='red', lwd=3)

```

## Convergence when n=2000, lambda = 100
```{r convergence3, include=TRUE, echo=FALSE, cache=TRUE}
lambda0 <- 2*10^(-6)
n <- 1000
k <- 3
simData <- simulate(n, k)

y <- simData$y
theta <- simData$y
D <- simData$D
eta <- matrix(D%*%theta)
M <- simData$M
cholM <- simData$cholM
tau <- 0.05
lambda <- lambda0*n^k/factorial(k)
step <- 1

max_iter <- 100000

multi_step <- spingarn_multi_step(theta, eta, y, D, cholM,
                                 lambda, tau, step, max_iter, k)

theta <- multi_step[[1]]
eta <- multi_step[[2]]
theta_last <- prox_f1(theta, y, tau)

plot(simData$x,simData$f,type='l',col='blue',
ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
points(simData$x,y, cex=.6)
lines(simData$x,theta_last,col='purple', lwd=3)

```