### 2017-06-23 Notes/timing_and_convergence_examples.Rmd
added RcppEigen functions and some timing comparisons

---
title: "detrendr examples"
output: pdf_document
---

```{r load, include=FALSE, echo=FALSE, cache=FALSE}
library(Matrix)
library(devtools)
library(roxygen2)
library(Rcpp)
library(microbenchmark)
load_all("../Code/detrendr")

```

# Compare cholesky decompositions and solvers
```{r cholesky_comparison}
nSeq <- seq(500, 20000, 500) 
k <- 4

timing <- data.frame(n=nSeq, Matrix_chol = NA, Matrix_solve = NA, 
                     Eigen_chol = NA, chol_solve = NA)
i <- 1
for (n in nSeq){
  D <- get_Dk(n, k)
  M <-  as(Diagonal(n) + Matrix::crossprod(D), "dgCMatrix")
  theta <- rnorm(n)

  timing$Matrix_chol[i] <- median(microbenchmark(cholM <- chol(M))$time*1e-3)
  timing$Eigen_chol[i] <- median(microbenchmark(cholM2 <- chol_eigen(M))$time*1e-3)
  timing$Matrix_solve[i] <- median(microbenchmark(
    x1 <- as.numeric(Matrix::solve(cholM, Matrix::solve(t(cholM), theta))))$time*1e-3)
  timing$chol_solve[i] <- median(microbenchmark(
    x2 <- chol_solve(cholM, chol_solve(t(cholM), theta, k, FALSE), 
                   k, TRUE))$time*1e-3)
  i <- i+1
}

plot(Matrix_chol~n, timing, ylab="microseconds", 
     main="Matrix package sparse cholesky decomposition")
plot(chol_solve~n, timing, ylab ="microseconds", 
     main="RcppEigen sparse cholesky decomposition")
plot(chol_solve~n, timing, ylab = "microseconds", 
     main="My banded cholesky solver (black) vs. Matrix solver (blue)")
points(Matrix_solve~n, timing, col="blue")

```

## Some examples of baseline fitting with detrendr

```{r simulate, include=FALSE, echo=FALSE}

simulate <- function(n, k){
  x <- seq(1/n, 1, length.out=n)
  f <- 2*(x + 2)^2 + 3*cos(3*pi*x)
  tau <- 5e4
  g1 <- 40*exp(-tau*(x-0.3)^2)
  g2 <- 30*exp(-tau/2*(x-0.58)^2)
  g4 <- 45*exp(-tau/7*(x-0.6)^2)
  g3 <- 37*exp(-tau*(x-0.63)^2)
  y <- f + g1 + g2 + g3 + g4 + rnorm(n)
  D <- get_Dk(n, k)
  M <- diag(n) + crossprod(D)
  cholM <- as.matrix(chol(M))
  return(list(y=y, f=f, x=x, D=D, M=M, cholM=cholM))
}
```

<!-- ## Current timing experiments -->
<!-- ```{r timing, include=FALSE, echo=FALSE, cache=TRUE, } -->
<!-- nSeq <- seq(100, 1500, 200) -->
<!-- timing <- data.frame(n=nSeq, R = NA, C=NA) -->
<!-- tau <- 0.01 -->
<!-- lambda <- 10 -->
<!-- step <- 1 -->

<!-- for (i in 1:length(nSeq)){ -->
<!--   simData <- simulate(nSeq[i], 3) -->
<!--   y <- simData$y -->
<!--   theta <- simData$y -->
<!--   D <- simData$D -->
<!--   eta <- matrix(D%*%theta) -->
<!--   M <- simData$M -->
<!--   cholM <- simData$cholM -->
<!--   timing$R[i] <- system.time(  -->
<!--       multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda,  -->
<!--                                             tau, step, 100))[3] -->
<!--   timing$C[i] <- system.time( -->
<!--     multi_step <- spingarn_multi_step(theta, eta, y, D, cholM, -->
<!--                                      lambda, tau, step, 100))[3] -->
<!-- } -->

<!-- plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue") -->
<!-- lines(R~n, timing, type="l") -->
<!-- ``` -->

<!-- ## Convergence when n=100, lambda = 10 -->
<!-- ```{r convergence, include=TRUE, echo=TRUE, cache=TRUE} -->

<!-- n <- 500 -->
<!-- simData <- simulate(n, 3) -->

<!-- y <- simData$y -->
<!-- theta <- simData$y -->
<!-- D <- simData$D -->
<!-- eta <- matrix(D%*%theta) -->
<!-- M <- simData$M -->
<!-- cholM <- simData$cholM -->
<!-- tau <- 0.05 -->
<!-- lambda <- 10 -->
<!-- step <- 1 -->

<!-- max_iter <- 70000 -->
<!-- rerr <- double(max_iter-1) -->
<!-- THx <- matrix(NA, n, max_iter) -->
<!-- for (iter in 1:max_iter) { -->
<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
<!--   theta <- one_step[[1]] -->
<!--   eta <- one_step[[2]] -->
<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
<!--   if (iter > 1){ -->
<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
<!--   } -->
<!-- } -->
<!-- theta_last <- prox_f1(theta, y, tau) -->

<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
<!-- plot(log(rerr), type="l") -->

<!-- plot(simData$x,simData$f,type='l',col='blue', -->
<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
<!-- points(simData$x,y, cex=.6) -->
<!-- lines(simData$x,theta_last,col='red', lwd=3) -->

<!-- ``` -->

<!-- ## Convergence when n=1000, lambda = 100 -->
<!-- ```{r convergence2, include=TRUE, echo=FALSE, cache=TRUE} -->

<!-- n <- 1000 -->
<!-- simData <- simulate(n, 3) -->

<!-- y <- simData$y -->
<!-- theta <- simData$y -->
<!-- D <- simData$D -->
<!-- eta <- matrix(D%*%theta) -->
<!-- M <- simData$M -->
<!-- cholM <- simData$cholM -->
<!-- tau <- 0.05 -->
<!-- lambda <- 100 -->
<!-- step <- 1 -->

<!-- max_iter <- 70000 -->
<!-- rerr <- double(max_iter-1) -->
<!-- THx <- matrix(NA, n, max_iter) -->
<!-- for (iter in 1:max_iter) { -->
<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
<!--   theta <- one_step[[1]] -->
<!--   eta <- one_step[[2]] -->
<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
<!--   if (iter > 1){ -->
<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
<!--   } -->
<!-- } -->
<!-- theta_last <- prox_f1(theta, y, tau) -->

<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
<!-- plot(log(rerr), type="l") -->

<!-- plot(simData$x,simData$f,type='l',col='blue', -->
<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
<!-- points(simData$x,y, cex=.6) -->
<!-- lines(simData$x,theta_last,col='red', lwd=3) -->

<!-- ``` -->

<!-- ## Convergence when n=2000, lambda = 100 -->
<!-- ```{r convergence3, include=TRUE, echo=FALSE, cache=TRUE} -->

<!-- n <- 2000 -->
<!-- simData <- simulate(n, 3) -->

<!-- y <- simData$y -->
<!-- theta <- simData$y -->
<!-- D <- simData$D -->
<!-- eta <- matrix(D%*%theta) -->
<!-- M <- simData$M -->
<!-- cholM <- simData$cholM -->
<!-- tau <- 0.05 -->
<!-- lambda <- 100 -->
<!-- step <- 1 -->

<!-- max_iter <- 70000 -->
<!-- rerr <- double(max_iter-1) -->
<!-- THx <- matrix(NA, n, max_iter) -->
<!-- for (iter in 1:max_iter) { -->
<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
<!--   theta <- one_step[[1]] -->
<!--   eta <- one_step[[2]] -->
<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
<!--   if (iter > 1){ -->
<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
<!--   } -->
<!-- } -->
<!-- theta_last <- prox_f1(theta, y, tau) -->

<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
<!-- plot(log(rerr), type="l") -->

<!-- plot(simData$x,simData$f,type='l',col='blue', -->
<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
<!-- points(simData$x,y, cex=.6) -->
<!-- lines(simData$x,theta_last,col='red', lwd=3) -->

<!-- ``` -->