### 2018-10-01 Notes/100118_meeting_notes.Rmd
reorganizing

---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(devtools)
require(Matrix)
require(ggplot2)
load_all("../detrendify/Code/detrendr")
```

## R Markdown

```{r dataimport}
dataDir <- "../../Desktop/SPod_Data/TestRange_Dec2017"
datafiles <- dir(dataDir, pattern=".csv", full.names=TRUE)
spod <- read.csv(datafiles[3], header=TRUE,  na.strings = "N/A")
spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
                                 format= "%m/%d/%Y %H:%M:%S"))

node <- "f"
pidCol <- paste(node, "SPOD.PID..V.", sep=".")
spodNode <- spod[, c("time", pidCol)]
names(spodNode)[2] <- c("pid")
spodNode <- subset(spodNode, !is.na(pid))
plot(pid~time, spodNode, type="l", main = "raw")

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.