### 2018-10-26 Notes/Gurobi_notes.Rmd
Meeting notes updated

---
title: "detrendr Meeting notes"
output: html_document
---

```{r setup, include=FALSE}
library(gurobi)
library(plyr)
library(devtools)
library(microbenchmark)
library(trustOptim)
library(Rglpk)
library(tidyverse)
library(parallel)
library(lubridate)
library(lbfgs)
load_all("../Code/detrendr")
```

```{r load, include = FALSE}
dataDir <- "~/Desktop/EPA/SPod_Data/TestRange_Dec2017"
datafiles <- dir(dataDir, pattern=".csv", full.names=TRUE)
spod <- read.csv(datafiles[3], header=TRUE,  na.strings = "N/A")
spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
                                 format= "%m/%d/%Y %H:%M:%S"))

node <- "h"
pidCol <- paste(node, "SPOD.PID..V.", sep=".")
spodNode <- spod[, c("time", pidCol)]
names(spodNode)[2] <- c("pid")
spodNode <- subset(spodNode, !is.na(pid))
spodNode$pid <- as.numeric(scale(spodNode$pid, center = TRUE))
```

## Gurobi vs Rglpk Timing Experiments

```{r, cache = TRUE}
tau <- 0.05
k <- 3

m_gurobi <- microbenchmark(
   gurobi_trend(spodNode$pid[1:200], tau, 200, k),
   gurobi_trend(spodNode$pid[1:400], tau, 400, k),
   gurobi_trend(spodNode$pid[1:1000], tau, 1000, k),
   gurobi_trend(spodNode$pid[1:2000], tau, 2000, k),
   times = 10
)

m_Rglpk <- microbenchmark(
  lpglpk_trendfilter(spodNode$pid[1:200], tau, 200, k),
  lpglpk_trendfilter(spodNode$pid[1:400], tau, 400, k),
  lpglpk_trendfilter(spodNode$pid[1:1000], tau, 1000, k),
  lpglpk_trendfilter(spodNode$pid[1:2000], tau, 2000, k),
  times = 10
)

knitr::kable(data.frame(n = c(200, 400, 1000, 2000), gurobi = summary(m_gurobi)$mean, Rglpk = summary(m_Rglpk)$mean), digits = 0)
```

### Approximation to objective function
Trend filtering objective:
min $\sum_i\rho_{\tau}(y_i - \theta_i) + ||D\theta||_1$

Approximate objective: 
Define $\hat{f}(x, \tau) = \sum_i \frac{1}{2*10^4\pi}log[cosh(10^4\pi x_i)] + (\tau - \frac{1}{2}) x_i$

min $\hat{f}(y - \theta, \tau) + \hat{f}(D\theta, 0.5)$

Both gradient and Hessian are easy to compute. Hessian is banded and sparse.

## Compare times on n=200
```{r, cache = TRUE}
lambda <- 200/2
y <- spodNode$pid[1:200]
theta_g <- gurobi_trend(y, tau, 200, k)
D <- get_Dk(length(y), k)

phi0 <- y - quantile(y, tau)

m2 <- microbenchmark(
  theta_g <- gurobi_trend(y, tau, 200, k),
  lbfgs_sol <- lbfgs(obj, grad_obj, vars = phi0,  y=y, tau=tau, D=D, 
                  lambda = lambda, invisible = 1),
  trust_sol <- trust.optim(phi0, obj, grad_obj, hess_obj, 
                        method = "Sparse", 
                        control = list(report.level = 0), 
                        y=y, tau=tau, 
                        D=D, lambda=lambda),
  times = 5
)

theta_lbfgs <- y-lbfgs_sol$par
theta_trust <- y-trust_sol$solution

knitr::kable(data.frame(method = c("Gurobi", "LBFGS", "Trust Region"), 
                        time = summary(m2)$mean), digits = 1)
```
## Compare results

```{r}
plot(y, type="l", col="grey")
lines(theta_g, col="red")
lines(theta_lbfgs, col="blue")
lines(theta_trust, col = "darkgreen")
```

## Cross validation to choose lambda - simulated data

```{r, cache = TRUE, echo = TRUE, message = FALSE, results=FALSE}
n <- 400
x <- seq(1, n, 1)
y <- sin(x*2*pi/n) + rnorm(n, 0, .5)

k <- 3
tau <- c(0.05, 0.5)

lambdaSeq <- seq(50, 1600, 50)
loss_cv <- lambda_cv(y, tau, k, lambdaSeq,
                     numFolds = 5, parallel = TRUE)
plot(rowSums(loss_cv$loss)~lambdaSeq, 
     ylab = "Check loss of validation data", 
     xlab = "Lambda")

theta_df <- get_windows(y, x, k, tau, loss_cv$lambda, n, 0)
plot.df <- left_join(theta_df, 
                     data.frame(time=x, y=y, 
                                true = sin(x*2*pi/n))) 
```

```{r, echo = FALSE}
# Chosen lambda
loss_cv$lambda

ggplot(plot.df, aes(x = time, y = y)) +
  geom_line(alpha = 0.2) +
  geom_line(aes(y=true), col="blue")+
  geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
  theme_bw() + 
  guides(col = "none")

```

## Cross-validation on real data

```{r, cache = TRUE}
k <- 3
tau <- c(0.05, 0.5)
df <- spodNode
y <- df$pid[1:3600]
n <- length(y)
lambdaSeq <- c(10*n, 12*n, 14*n, 16*n)

spod_cv <- lambda_cv(y, tau, k, lambdaSeq,
          numFolds = 5, parallel = TRUE)
plot(rowSums(spod_cv$loss)~lambdaSeq, 
     ylab = "Check loss of validation data", 
     xlab = "Lambda")
spod_cv$lambda
```

```{r, include=FALSE}
theta_df <- get_windows(y, df$time[1:n], k, tau, loss_cv$lambda, n, 0)
plot.df <- left_join(theta_df, spodNode) 
```

```{r, echo = FALSE, message = FALSE}
ggplot(plot.df, aes(x = time, y = pid)) +
  geom_line(alpha = 0.2) +
  geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
  theme_bw() + 
  guides(col = "none")
```

# Complete day using hour long windows 
```{r, eval= TRUE, cache = TRUE, include = FALSE}
theta_df <- get_windows(df$pid, df$time, k, tau, spod_cv$lambda, 
                        length(y), 600)
plot.df <- left_join(theta_df, spodNode) 
```

```{r, message = FALSE}
ggplot(plot.df, aes(x = time, y = pid)) +
    geom_line(alpha = 0.2) +
    geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
    theme_bw()
```

```{r, message=FALSE}
ggplot(plot.df, aes(x = time, y = pid)) +
    geom_line(alpha = 0.2) +
    geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
    xlim(c(ymd_hms("2017-11-30 9:00:00", tz = "EST"), 
           ymd_hms("2017-11-30 11:30:00", tz = "EST")))+
    theme_bw()
```

# Compare all 3 nodes

```{r, message = FALSE}
spod_pid <- spod %>%
  select(time, contains("SPOD.PID..V.")) %>%
  mutate_at(vars(-time), function(x) as.numeric(scale(x, center=TRUE))) %>%
  gather("node", "pid", -time) %>%
  mutate(node_name = substr(node, 1, 1))

ggplot(spod_pid, aes(x=time, y=pid, col=node_name)) + 
  geom_line(alpha = 0.2) + 
  scale_color_brewer(palette = "Set1") + 
  theme_bw()
```

## Use selected lambda to detrend

```{r, cache = TRUE, message = FALSE}
nodes <- c("f", "g", "h")
plot_df <- data.frame(time = spodNode$time)
tau <- 0.05
for (node in nodes){
  pidCol <- paste(node, "SPOD.PID..V.", sep=".")
  spodNode <- spod[, c("time", pidCol)]
  names(spodNode)[2] <- c("pid")
  spodNode$pid <- as.numeric(scale(spodNode$pid, center = TRUE))
  theta_df <- get_windows(spodNode$pid, spodNode$time, k, tau,
                          spod_cv$lambda, 3600, 600)
  theta_comb <- theta_df %>% 
    group_by(time) %>% 
    summarise(
      theta = mean(theta)
      )

  spodNode[,sprintf("pidC_%s", node)] <- spodNode$pid - theta_comb$theta
  plot_df <- left_join(plot_df, spodNode[,c(1,3)]) 
}

plot_long <- plot_df %>% gather("node", "pid", -time)

ggplot(plot_long, aes(x=time, y=pid, col=node)) + 
  geom_line(alpha = 0.2) + 
  scale_color_brewer(palette = "Set1") + 
  theme_bw()

ggplot(plot_long, aes(x=time, y=pid, col=node)) + 
  geom_line(alpha = 0.3) + 
  scale_color_brewer(palette = "Set1") + 
  xlim(c(ymd_hms("2017-11-30 9:30:00", tz = "EST"), 
         ymd_hms("2017-11-30 11:00:00", tz = "EST")))+
  theme_bw()
```