### 2019-01-24 Manuscript/application_short.tex
manuscript and figure edits

\documentclass[12pt]{article}
\RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
\usepackage{graphicx}
%\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage{placeins}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

	\section{Method comparison application}
	\begin{figure}[!h]
		\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile and a threshold of the median plus 3MAD.}
		\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\caption{Variation of Information (VI) between sensor nodes after trend removal by quantile and method. The rows in the figure represent the three thresholds used: 3MAD, 4MAD, and 5MAD. Lower VI represents more similar classifications. }
		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
	\end{figure}

		\begin{figure}[!h]
		\centering
		\caption{Normalized Mutual Information (NMI) between sensor nodes after trend removal by quantile and method. The rows in the figure represent the three thresholds used: 3MAD, 4MAD, and 5MAD. Higher NMI represents more similar classifications. }
		\includegraphics[width = .9\linewidth]{Figures/NMI_app_short.png}
	\end{figure}

	\newpage
	\input{short_confusion_detrend_MAD3.tex} 
	\input{short_confusion_detrend_MAD4.tex} 
	\input{short_confusion_detrend_MAD5.tex}

\end{document}